---
title: "10x scRNA-seq Seurat Analysis"
author: "PRIME (by2747@stowers.org)"
output:
  html_document:
    code_folding: hide
    prettydoc::html_pretty:
      theme: leonids
      toc: yes
params:
  Reference_mito_list: null
---

``` {r setup, echo=FALSE, message=FALSE, results="hide", warning=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(pander)
library(knitr)
library(rmarkdown)
library(ggplot2)

library(rmdformats)
library(prettydoc)
library(hrbrthemes)
library(tint)
library(tufte)

panderOptions('table.style','rmarkdown')
panderOptions('table.split.table',Inf)


options(width=50)
opts_chunk$set(error=FALSE)
opts_chunk$set(echo=FALSE)
opts_chunk$set(tidy=FALSE)
opts_chunk$set(warnings=FALSE)
```

## Method 
* [Cellranger](https://www.10xgenomics.com/support/software/cell-ranger/latest) count version 8.0.1 was used in the initial single cell analysis of the samples. </br>
* [CellBender](https://www.nature.com/articles/s41592-023-01943-7) version 0.3.0 were used on h5 files output by Cellranger for further denoising. </br>
* [Seurat](https://satijalab.org/seurat/) version 5.1.0 was used for data filteration, normalization, and single cell analysis. </br>


## Data QC
</br>

Feature violin plots 
```{r QCplot-violin, echo=TRUE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6,fig.path='Seurat_Report_gen_plots/'}
SC_cellbender_OutDir <- "./"
SC_10X_cellBenderH5.filteredSeurat <- paste0(SC_cellbender_OutDir, "cellbender_output_filtered_seurat.h5")

SC_10X.data <- Read10X_h5(SC_10X_cellBenderH5.filteredSeurat)
SC_10X <- CreateSeuratObject(counts = SC_10X.data, project = "10X", min.cells = 3, min.features = 200)
num_cells_preFilter <- length(SC_10X@meta.data$nCount_RNA)

#Check the mt gene list and select mt features based on said list
mt_file <- params$Reference_mito_list
if (is.null(mt_file)) {
  stop("Error: The mt gene list is NULL. Stopping the rmarkdown process.")
}
mt_gene_list <- readLines(mt_file)
mt_gene_list_inter <- intersect(mt_gene_list, rownames(SC_10X)) #Make sure that mt genes are present in the SC obj
SC_10X[["percent.mt"]] <- PercentageFeatureSet(SC_10X, features = mt_gene_list_inter)

#Visualize QC metrics as a violin plot and scatter plot
VlnPlot(SC_10X, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

Feature scatter plot
```{r QCplot-scatter, echo=TRUE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6, fig.path='Seurat_Report_gen_plots/'}
plot1 <- FeatureScatter(SC_10X, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(SC_10X, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

## Data filtering and normalization 
Cells with nFeature_RNA greater than 200 and with less than 10% mitochondria counts were selected for further processing. Code can be visualized below: 

```{r filtering, echo=TRUE, warning=FALSE, message=FALSE }
SC_10X <- subset(SC_10X, subset = nFeature_RNA > 200 & percent.mt < 10)
```

After removing unwanted cells from the dataset, the next step is to normalize the data. SCTransform was used here for data normalization. Code can be visualized below:

```{r nomorlize, echo=TRUE, warning=FALSE, message=FALSE }
# SC_10X <- NormalizeData(SC_10X, normalization.method = "LogNormalize", scale.factor = 10000)
SC_10X <- SCTransform(SC_10X, vars.to.regress = "percent.mt", verbose = FALSE)
num_cells_postFilter <- length(SC_10X@meta.data$nCount_RNA)

```

*After data filtering and normalization, the total number of cells within the dataset changed from `r num_cells_preFilter` to `r num_cells_postFilter`* </br>

## Cluster Analysis

To help determine the ‘dimensionality’ of the dataset, we calcualted the PCA of the data, and from the results generated an ‘Elbow plot’. 
ndim of 30 was picked for further analysis (red line), as we believe this value can better explain the majority of the data.  

```{r dimension-elbow, echo=TRUE, warning=FALSE, message=FALSE, fig.width=11, fig.height=6, fig.path='Seurat_Report_gen_plots/'}

SC_10X <- RunPCA(SC_10X, verbose = FALSE, npcs=100)

# Create Elbow plot using ggplot2
ElbowPlot(SC_10X, ndims = 100) + geom_vline(xintercept=30,color="red")

```
```{r BioTools-BiotoolsElbow, echo=FALSE, warning=FALSE, message=FALSE}
plot_U <- ElbowPlot(SC_10X, ndims = 80) + theme(axis.title=element_text(size=17,face="bold"), axis.text=element_text(size=15)) + geom_vline(xintercept=30,color="red")
ggsave(filename = "Seurat_Report_gen_plots/BioTool-dimension-elbow.png", plot = plot_U, width = 8, height = 8, dpi = 500 )
```

Next we perform UMAP embedding on the normalized (SCTransform) data. A dimPlot is shown below for reference using dim 1:30, and a default resolution of 0.5. 

```{r scalling_PCA, echo=TRUE, warning=FALSE, message=FALSE, fig.width=11, fig.height=8, fig.path='Seurat_Report_gen_plots/'}
all.genes <- rownames(SC_10X)
# SC_10X <- ScaleData(SC_10X, features = all.genes)

SC_10X <- RunUMAP(SC_10X, dims = 1:30, verbose = FALSE)

SC_10X <- FindNeighbors(SC_10X, dims = 1:30, verbose = FALSE)
SC_10X <- FindClusters(SC_10X, verbose = FALSE, resolution = 0.5) 

DimPlot(SC_10X, label = TRUE)
```

```{r BioTools-BiotoolsUMAP, echo=FALSE, warning=FALSE, message=FALSE}
plot_U <- DimPlot(SC_10X, label = TRUE) + theme(axis.title=element_text(size=17,face="bold"), axis.text=element_text(size=15))
ggsave(filename = "Seurat_Report_gen_plots/BioTool-UMAP.png", plot = plot_U, width = 8, height = 8, dpi = 500 )
```

## Find Marker genes

Seurat can help you find markers that define clusters via differential expression (DE). In this initial analysis, we chose to identify positive markers - genes that are expressed higher in a given cluster than in the rest of the data. </br>

```{r findMarkers, echo=TRUE, warning=FALSE, message=FALSE}
SC_10X.markers <- FindAllMarkers(SC_10X, only.pos = TRUE, logfc.threshold = 0.5)
SC_10X.markers <- SC_10X.markers %>% filter(p_val_adj < 0.1)

write.csv(SC_10X.markers, 'Seurat_Report_gen_plots/AllMarkers.csv')
cwd <- getwd()
url <- paste0("https://webfs/", cwd, "/../PRIME_SC_out/Seurat_Report_gen_plots/AllMarkers.csv")
```
Using DoHeatmap(), we are plotting the top 5 markers (or all markers if less than 5) for each cluster (log2FC > 0.5 & adjusted Pvalue < 0.1  were applied for marker selection).</br>
</br>
Table for all marker genes can be downloaded [AllMarkers.csv](`r url`)

```{r markers-heatmap, echo=TRUE, warning=FALSE, message=FALSE, fig.width=10, fig.height=12, fig.path='Seurat_Report_gen_plots/'}
SC_10X.markers %>%
    group_by(cluster) %>%
    # dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top5
DoHeatmap(SC_10X, features = top5$gene) + NoLegend()
```
We also plotted the markers above into a dotplot using DotPlot()
```{r markers-dotplot, echo=TRUE, warning=FALSE, message=FALSE, fig.width=14, fig.height=12, fig.path='Seurat_Report_gen_plots/'}
features <- unique(top5$gene) # Dotplot can not have duplicates in the levels of features

DotPlot(object = SC_10X, features = features) +
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ylab("Cluster") +
  coord_flip()
  
url <- paste0("https://webfs/", cwd, "/../PRIME_SC_out/SeuratObj.rds")
PRIME_SC_out_folder <- paste0("https://webfs/", cwd, "/../PRIME_SC_out")
```

## Data Download
**Seurat object saved to Prime_SC_Analysis/ folder under [SeuratObj.rds](`r url`)** </br>
**Additional Seurat analysis outputs can be found in this folder here [PRIME_SC_out](`r PRIME_SC_out_folder`)**

```{r save, echo=TRUE, warning=FALSE, message=FALSE }
saveRDS(SC_10X, file = "SeuratObj.rds")
``` 

###  Session information

**Author:** PRIME **Bowen Yang (by2747@stowers.org)******

**Contact:** Please email helpCompbio@stowers.org if you have any questions or suggestions.

**Generated:** `r format(Sys.time(), "%a %b %d %Y, %I:%M %p")`

For reproducibility, see session info below:

``` {r session_info, echo=FALSE, comment=NA}
sessionInfo()
```
